Eleven years of student replication projects provide evidence on the correlates of replicability in psychology

### 2023-04-05 paper/manuscript.Rmd
more intro

# TODO list
- BETTER abstract
- many citations everywhere
- long list of things to gesture at in intro
- link to open stuffs
- ** verify code for ES ** with someone
- redo intro
- redo results
- redo discussion
- fill in todos

@@

@@

@wagge2019 get UGs to do replications for

@quintana2021 do reps as honors thesis
@pownall2021 sips 2021 on teaching openness
@jekel2020 teach open science w/ replication 
@jern2018 in one study didn't show critical thinking benefits for students to make them do replications

# introduction
[wacky metaphor, maybe drop] When building a stone wall, it's important that the components be stones that will retain their shape and not clots of dirt and sand that might fall apart and bring the rest of the wall falling down. Building a cumulative body of scientific theories similarly requires that the empirical results, especially the more foundational empirical results, be sturdy. Results that cannot be observed consistently are like dirt clots, they don't hold their shape and thus are not a strong foundation for the theories and studies built on top of them. Replications and the theory of replicability can be thought of as a way of testing whether the empirical results are sturdy. Replicating is a necessary feature for load-bearing empirical results. We note that replicability is not the only desirable feature of a study, and that replicating does not mean that a study is valid, or generalizable, or interesting.

@@

As an alternative, much of the work looking at predictors of replicability has been correlation: whether using regression models, prediction markets, or machine learning techniques to build predictions about what will replicate. However, all of these approaches depend on data from replications. While many replications have been conducted, most were not randomly sampled and are not pooled together, so these studies instead rely on the small number of data points from large-scale replication projects. The RP:P dataset itself is much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016;@anderson2016].

@@

Our main contribution is to add a collection of 176 student replication projects to the available pool. We use these studies to examine the correlates of replicability.

our contribution: provide more data!
using pedagogical replications, which are good because they’re p(build on)
slightly different estimand but maybe a better one?

## predictors
In addition to determining estimated overall replicaiton rates for fields and journals, there's also merit in knowing what features of experiments (and replication attempts) are predictive of replication success. Blah blah stakeholders and resources. Scientists may want to build their work on studies that are likely to replicate, so that they can replicate and build on work without wasting resources. Similarly investing in interventions may want to start with things that are more likely to replicate. There's some signal here, as people are able to predict replication success at above chance CITATIONS [@hoogeveen2019;@dreber2015; @forsell2019;@camerer2018]

RP:P looked at how replication rates varied across subsamples of their studies [@openscienceconsortium2015]. They found that cognitive psychology studies replicated at higher, but still low rates (50% v 25%) compared to social psychology. They also found that larger effect sizes and smaller p-values of original studies were predictive of replicating. TODO do we talk about any other correlates.

## current work

Prior approaches to replicability have focused on a potentially problematic estimand: the probability of a finding in the literature being somehow truly replicable. Critics have pointed out that "true" replicability may not be possible to estimate outside of a specific sample (CITE van bavel etc.) or even time period (CITE ramscar).

Further, the methods for estimating this quantity have been theoretically problematic. Sampling schemes for prior work typically does not reflect a random sample from the literature -- and even if they did, arguably a literature will succeed if useful discoveries come out of it, not if random findings are true ("Science is not signal detection" ref).

In contrast, we have pursued a different estimand: the probability that a researcher, on selecting a finding of interest from the literature, can successfully achieve a result satisfactorily close enough to the original that they feel that they can build on it in their own work, with all the necessary compromises to the methods and sample of the original that may be required by the constraints of the situation.

There's different approaches one could take to sampling, we take the versimilitude approach. Our sample is not a random sample of the literature, or of a given journal, or some weighting, but it does represent studies that people are interested in replicating.

Here we introduce a new dataset of 176 replications of experimental studies from the social sciences, primarily psychology. These replications were conducted by students in graduate-level experimental methods class between 2011 and 2022 as individual course projects. This approximately doubles the set of experiments in the large-scale replication literature. We investigate predictors of replicability in this new dataset.

@@

TODO discuss more about why this one and what other measure we don't choose are

We used a subjective rating of replication success as our primary outcome measure. The instruction team had coded a holistic measure of replication success for each project when they were turned in at the end of the course. For reliability, VB independently code the replication success from the replication reports; discrepancies were resolved by discussion between MCF and VB (`r pluck(matching,2,1)/176*100 |> round()` % of cases). This measure was applicable across the range of statistical measures and reporting practices and accommodated studies where there were multiple important outcome measures.

As a complement, we also used two statistical measure of replication on the subset of the data where they were computable (`r pluck(pred_int_avail,1)` cases, see Figure \ref{fig:prisma}). We measured p-original, the p-value on the null hypothesis that the original and replication statistics are from the same distribution, as a continuous variable, and we also determined whether the replication statistic fell within the prediction interval of the original statistic [@errington2021].

@@

Distributions of study outcomes across some of these properties are shown in Figure \ref{fig:predictors-graph}. Both social and cognitive psychology studies were well represented, and the cognitive psychology studies replicated at twice the rate of social psychology studies. Within and between subjects designs were both common, and within replicated four times as much. Similarly, studies with multiple vignettes replicated 1.5 times more than single vignetted studies. However, there were strong correlations among these experimental features and between these experimental features and subfield.

@@

Results from other models can be found in the supplement. TODO make this a true statement

@@

In line with previous replications, we found a FOOBAR replication rate, with some studies showing effect sizes similar to the original and others much smaller. Some individual correlates of replicability that stood out included within-subjects designs, work in the subfield of cognitive psychology, and the original and replication both using online samples. TODO other possible things

Many of these correlates interrelate with one another, and we are still limited in our sample, so models with multiple predictors have a lot of uncertainty around the effects of each predictor.

>>>>

<!--
# TODO list
- BETTER abstract
- many citations everywhere
- long list of things to gesture at in intro
- link to open stuffs
- ** verify code for ES ** with someone
- redo intro
- redo results
- redo discussion
- fill in todos

@@

## predictors
In addition to determining estimated overall replicaiton rates for fields and journals, there's also merit in knowing what features of experiments (and replication attempts) are predictive of replication success. Blah blah stakeholders and resources. Scientists may want to build their work on studies that are likely to replicate, so that they can replicate and build on work without wasting resources. Similarly investing in interventions may want to start with things that are more likely to replicate. There's some signal here, as people are able to predict replication success at above chance CITATIONS

RP:P looked at how replication rates varied across subsamples of their studies [@openscienceconsortium2015].

using pedagogical replications, which are good because they’re p(build on)
slightly different estimand but maybe a better one?
One key part of the question of whether a study is replicable or trustworthy is the idea of "if I did this, would I get the same pattern of results?".

@@

@pownall2021 sips 2021 on teaching openness
@jekel2020 teach open science w/ replication 
@jern2018 in one study didn't show critical thinking benefits for students to make them do replications
-->
# introduction
When building a stone wall, it's important that the components be stones that will retain their shape and not clots of dirt and sand that might fall apart and bring the rest of the wall falling down. Building a cumulative body of scientific theories similarly requires that the empirical results, especially the more foundational empirical results, be sturdy. Results that cannot be observed consistently are like dirt clots, they don't hold their shape and thus are not a strong foundation for the theories and studies built on top of them. Replications and the theory of replicability can be thought of as a way of testing whether the empirical results are sturdy. Replicating is a necessary feature for load-bearing empirical results. We note that replicability is not the only desirable feature of a study, and that replicating does not mean that a study is valid, or generalizable, or interesting.

@@

As an alternative, much of the work looking at predictors of replicability has been correlation: whether using regression models, prediction markets, or machine learning techniques to build predictions about what will replicate. However, all of these approaches depend on data from replications. While many replications have been conducted, most were not randomly sampled and are not pooled together, so these studies instead rely on the small number of data points from large-scale replication projects. The RP:P dataset itself is much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016;@anderson2016]. These correlational approaches have established that people can predict what studies will replicate above chance [@hoogeveen2019;@dreber2015; @forsell2019;@camerer2018], but have not identified that many concrete predictors that differentiate replications from non-replications. In the RP:P sample, studies in cognitive psychology (as opposed to social psychology) and studies with larger effect sizes and smaller p-values were more likely to replicate.

@@

The perceived need for more replications in tension with the lack of publishing incentives for doing replications led several to suggest that pedagogical replications could fill this gap. Doing replications is a way for students to learn to do experiments in a scaffolded way and leads to more useful results than what students could design from scratch on the same timescale [@quintana2021; @wagge2019;@frank2012; @hawkins]. [TODO needs transition]

Our main contribution is a new dataset of 176 replications of experimental studies from the social sciences, primarily psychology. These replications were conducted by students in graduate-level experimental methods class between 2011 and 2022 as individual course projects. This approximately doubles the set of experiments in the large-scale replication literature. We investigate predictors of replicability in this new dataset.

Prior approaches to replicability have focused on interpreting results in terms of a potentially problematic estimand: the probability of a finding in the literature being somehow truly replicable. Critics have pointed out that "true" replicability may not be possible to estimate outside of a specific sample (CITE van bavel etc.) or even time period (CITE ramscar).

Further, the methods for estimating this quantity have been theoretically problematic. Sampling schemes for prior work typically do not reflect an entirely random sample from the literature; instead they sample from specific journals where results may be of more interest and adjust the sample for feasibility concerns. These are reasonable sampling choices, but they undermine the claim that the estimand is the level of "truth" in the literature as a whole. Sampling truely at random from the literature may not even be desirable, as  arguably a literature will succeed if useful discoveries come out of it, not if random findings are true (CITE "Science is not signal detection" ref). How much we care about whether a study is "trustworthy" is not uniformly distributed across the literature.

In contrast, we have explicitly pursued a different estimand: the probability that a researcher, on selecting a finding of interest from the literature, can successfully achieve a result satisfactorily close enough to the original that they can build on it in their own work, with all the necessary compromises to the methods and sample of the original that may be required by the constraints of the situation.

The sampling for our replications is based on what studies students were interested in and wanted to replicate, with some constraints of feasibility; this sampling reflects how scientists choose what studies to build on: those that are interesting and relatively doable within their ability and budgetary constraints. Our primary criteria for replication success is a subjective measure of whether the study replicated for the same reason, that we believe this is usually what researchers care about: is it close enough that you think you're getting at the same effect and could confidently build on it. The quality of the replications also reflects this: unlike some prior work like @camerer2018, which explicitly aimed for high-powered replications, the power in these studies was limited by budgetary reasons in some cases, which is reflective of the types of resources often available to trainee scientists.

We think this is a valuable estimand because it's estimating what people actually care about: is this result of interest sturdy enough to support cumulative science?

@@

Many different measures can be used to define replication success of an individual statistical result, and there is much discussion over what each measure represents and which make sense [@simonsohn2015; @gelman2018; @mathur2020]. Because we operationalized replicability as whether a study could be built upon, we used a subjective rating of replication success as our primary outcome measure. This measure has the benefit that it was applicable across the diverse range of statistical measures and reporting practices present in the sample. It, unlike statistical measures of replication, could easily accommodate studies where there were multiple important outcome measures that together defined the pattern of interest. A holistic measure of replication success had been coded for each project when it was turned in at the end of the class. For reliability, VB independently code the replication success from the replication reports; discrepancies were resolved by discussion between MCF and VB (`r pluck(matching,2,1)/176*100 |> round()` % of cases).

As a complement, we also used two statistical measure of replication on the subset of the data where they were computable (`r pluck(pred_int_avail,1)` cases, see Figure \ref{fig:prisma}). We measured p-original, the p-value on the null hypothesis that the original and replication statistics are from the same distribution, as a continuous variable, and we also determined whether the replication statistic fell within the prediction interval of the original statistic [@errington2021]. These both measure how similar the estimates from the two sets of data are.

@@

Distributions of study outcomes across some of these properties are shown in Figure \ref{fig:predictors-graph}. Both social and cognitive psychology studies were well represented, and the cognitive psychology studies replicated at twice the rate of social psychology studies. Within and between subjects designs were both common, and within replicated four times as much. Similarly, studies with multiple vignettes replicated 1.5 times more than single vignetted studies. However, there were strong correlations among these experimental features and between these experimental features and subfield. TODO actual numbers here!

@@

Results from other models can be found in the supplement.

@@

In line with previous replications, we found a FOOBAR replication rate, with some studies showing effect sizes similar to the original and others much smaller. Some individual correlates of replicability that stood out included within-subjects designs, work in the subfield of cognitive psychology, and the original and replication both using online samples.  Many of these correlates interrelate with one another, and we are still limited in our sample, so models with multiple predictors have a lot of uncertainty around the effects of each predictor.
